### 2024-06-12 conteudo/01-exemplo-normas-ime.tex
Add another part of the process

%!TeX root=../tese.tex
%("dica" para o editor de texto: este arquivo é parte de um documento maior)
% para saber mais: https://tex.stackexchange.com/q/78101

\en

\chapter{Background}

\input{conteudo/background/codeduplication}
\par
\input{conteudo/background/nlpsimilarity}
\par
\input{conteudo/background/refactor}
\par
\input{conteudo/background/linux}
\par