### 2024-06-12 conteudo/background/refactor.tex

\en

\section{Refactoring}

Fowler defines refactoring in his book as "a controlled technique for imporving the design of an existing code base. Its
essence is applying a series of small behavior-preserving transformations, each of which too small to be worth doing. However
the cumulative effect of each of these transformations is quite significant." \citep{refactorbook}. The elimination of detected 
code clone duplication is a refactoring task that necessitates the application of refactoring methods. 
Fowler presents multiples refactoring methods such as parameterize method, replace parameter with explicit methods and 
replace parameter with method \citep{refactorbook}. There is a briefly explanation of the methods used in our research below.

\subsection{Parameterize method}

TODO. I EXPECT TO FILL THIS SECTION AS THE WORK EVOLVES.