### 2024-06-14 conteudo/tool/functionalities.tex
Add more work, 14/06

\en

\section{Tool functionalites}
\label{subsec:func}