### 2024-07-17 conteudo/background/refactor.tex

\en

\section{Refactoring}

Fowler defines refactoring in his book as "a controlled technique for improving the design of an existing code base. 
Its essence is applying a series of small behavior-preserving transformations, each of which is too small to be worth 
doing. However, the cumulative effect of each of these transformations is quite significant." \citep{refactorbook}. 
The elimination of detected code clone duplication is a refactoring task that necessitates the 
application of refactoring methods. Fowler presents multiple refactoring methods such as the 
\textit{parameterize method}, the \textit{replace parameter with explicit methods} and 
the \textit{replace parameter with method} \citep{refactorbook}. 
There is a brief explanation of the methods used in our research below.

\subsection{Parameterize method}

TODO. I EXPECT TO FILL THIS SECTION AS THE WORK EVOLVES.