### 2024-07-30 conteudo/eval/eval.tex
Add initial tool eval

On this Chapter, we exposes our experiment results bout the proposed tool performance following
the methology described in Chapter \ref{cha:method}. Firstly, we shows the results obtained by our tool
in the BigCloneBench dataset\citep{bigclonebench}, which is the standard form of the formal literature 
\citep{litreview}. Secondly, we shows the results obtained by our tool on the empirical analyses. 
Finaly, we do a triangulation of data analysing our finds on the tool performance.

\subsection{Tool results on the BigCloneBench dataset}

\input{conteudo/eval/table_bigclone}

\subsection{Tool results on the empirical analyses}

\input{conteudo/eval/table_empirical}

The results were obtained by randomly selecting ten functions in the AMD Display driver for 
each similarity range and analyzing if they are duplications and which duplication type.

\subsection{Discussion}