### 2024-07-31 conteudo/apendices/emp-lista.tex
format structure and remove artifacts from example

\chapter{Random selected samples}

THIS APPENDIX IS FOR LIST THE RANDOM SELECTED FUNCTIONS PAIRS IN THE 
EMPIRICIAL ANALYSES OF THE TOOL. I will put it here later.