### 2024-08-03 conteudo/apendices/cem.tex
create appendix

\section{Function pairs with similarity between 99\% and 100\%}

\paragraph{Function pair 1 }

\

\textbf{File 1:} dc/gpio/dcn20/hw\_factory\_dcn20.c

\textbf{Function name 1:} define\_ddc\_registers

\textbf{File 2:}a dc/gpio/dcn21/hw\_factory\_dcn21.c

\textbf{Function name 2:} define\_ddc\_registers

\textbf{Similarity:} 100\%

\textbf{Duplication type:} Type-1

\paragraph{Function pair 2 }

\

\textbf{File 1:} dc/gpio/dcn10/hw\_factory\_dcn10.c

\textbf{Function name 1:} dal\_hw\_factory\_dcn10\_init

\textbf{File 2:} dc/gpio/dcn21/hw\_factory\_dcn21.c

\textbf{Function name 2:} dal\_hw\_factory\_dcn21\_init

\textbf{Similarity:} 99.10\%

\textbf{Duplication type:} Type-4

\paragraph{Function pair 3 }

\

\textbf{File 1:} dmub/src/dmub\_dcn32.c

\textbf{Function name 1:} dmub\_dcn32\_is\_supported

\textbf{File 2:} dmub/src/dmub\_dcn35.c

\textbf{Function name 2:} dmub\_dcn35\_is\_supported

\textbf{Similarity:} 100\%

\textbf{Duplication type:} Type-1

\paragraph{Function pair 4 }

\

\textbf{File 1:} dc/bios/dce112/command\_table\_helper2\_dce112.c

\textbf{Function name 1:} clock\_source\_id\_to\_atom\_phy\_clk\_src\_id

\textbf{File 2:} dc/bios/dce60/command\_table\_helper\_dce60.c

\textbf{Function name 2:} clock\_source\_id\_to\_atom\_phy\_clk\_src\_id

\textbf{Similarity:} 100\%

\textbf{Duplication type:} Type-1

\paragraph{Function pair 5 }

\

\textbf{File 1:} dc/resource/dcn314/dcn314\_resource.c

\textbf{Function name 1:} dcn31\_opp\_create

\textbf{File 2:} dc/resource/dcn315/dcn315\_resource.c

\textbf{Function name 2:} dcn31\_opp\_create

\textbf{Similarity:} 100\%

\textbf{Duplication type:} Type-1

\paragraph{Function pair 6 }

\

\textbf{File 1:} dc/resource/dcn314/dcn314\_resource.c

\textbf{Function name 1:} dcn31\_apg\_create

\textbf{File 2:} dc/resource/dcn316/dcn316\_resource.c

\textbf{Function name 2:} dcn31\_apg\_create

\textbf{Similarity:} 100\%

\textbf{Duplication type:} Type-1

\paragraph{Function pair 7 }

\

\textbf{File 1:} dmub/src/dmub\_dcn20.c

\textbf{Function name 1:} dmub\_dcn20\_set\_outbox0\_rptr

\textbf{File 2:} dmub/src/dmub\_dcn35.c

\textbf{Function name 2:} dmub\_dcn35\_set\_outbox0\_rptr

\textbf{Similarity:} 100\%

\textbf{Duplication type:} Type-1

\paragraph{Function pair 8 }

\

\textbf{File 1:} dc/resource/dcn301/dcn301\_resource.c

\textbf{Function name 1:} dcn301\_dpp\_destroy

\textbf{File 2:} dc/resource/dcn31/dcn31\_resource.c

\textbf{Function name 2:} dcn31\_dpp\_destroy

\textbf{Similarity:} 100\%

\textbf{Duplication type:} Type-1

\paragraph{Function pair 9 }

\

\textbf{File 1:} dc/resource/dce110/dce110\_resource.c

\textbf{Function name 1:} dce110\_transform\_destroy

\textbf{File 2:} dc/resource/dce80/dce80\_resource.c

\textbf{Function name 2:} dce80\_transform\_destroy

\textbf{Similarity:} 100\%

\textbf{Duplication type:} Type-1

\paragraph{Function pair 10 }

\

\textbf{File 1:} dc/resource/dcn32/dcn32\_resource.c

\textbf{Function name 1:} dcn32\_clock\_source\_create

\textbf{File 2:} dc/resource/dcn321/dcn321\_resource.c

\textbf{Function name 2:} dcn321\_clock\_source\_create

\textbf{Similarity:} 100\%

\textbf{Duplication type:} Type-1