### 2024-08-03 conteudo/apendices/cinquenta.tex

\section{Function pairs with similarity between 49\% and 51\%}

\paragraph{Function pair 1 }

\

\textbf{File 1:} dc/hwss/dcn10/dcn10\_hwseq.c

\textbf{Function name 1:} dcn10\_init\_pipes

\textbf{File 2:} dc/hwss/dcn20/dcn20\_hwseq.c

\textbf{Function name 2:} dcn20\_program\_front\_end\_for\_ctx

\textbf{Similarity:} 50.27\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 2 }

\

\textbf{File 1:} dc/bios/command\_table\_helper.c

\textbf{Function name 1:} dal\_cmd\_table\_helper\_encoder\_id\_to\_atom

\textbf{File 2:} amdgpu\_dm/amdgpu\_dm.c

\textbf{Function name 2:} to\_drm\_connector\_type

\textbf{Similarity:} 50.51\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 3 }

\

\textbf{File 1:} dmub/src/dmub\_dcn32.c

\textbf{Function name 1:} dmub\_dcn32\_get\_inbox1\_wptr

\textbf{File 2:} dmub/src/dmub\_dcn31.c

\textbf{Function name 2:} dmub\_dcn31\_set\_inbox1\_wptr

\textbf{Similarity:} 49.64\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 4 }

\

\textbf{File 1:} dc/dcn30/dcn30\_mpc.c

\textbf{Function name 1:} mpc3\_program\_luta

\textbf{File 2:} dc/dpp/dcn10/dcn10\_dpp\_cm.c

\textbf{Function name 2:} dpp1\_cm\_program\_regamma\_lutb\_settings

\textbf{Similarity:} 50.96\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 5 }

\

\textbf{File 1:} dc/pg/dcn35/dcn35\_pg\_cntl.c

\textbf{Function name 1:} pg\_cntl35\_io\_clk\_status

\textbf{File 2:} dc/pg/dcn35/dcn35\_pg\_cntl.c

\textbf{Function name 2:} dcn\_pg\_cntl\_destroy

\textbf{Similarity:} 49.52\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 6 }

\

\textbf{File 1:} dc/dml/dcn314/display\_rq\_dlg\_calc\_314.c

\textbf{Function name 1:} is\_dual\_plane

\textbf{File 2:} dc/hubbub/dcn30/dcn30\_hubbub.c

\textbf{Function name 2:} convert\_and\_clamp

\textbf{Similarity:} 49.20\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 7 }

\

\textbf{File 1:} dc/dce/dce\_stream\_encoder.c

\textbf{Function name 1:} dce110\_stream\_encoder\_stop\_hdmi\_info\_packets

\textbf{File 2:} dc/dio/dcn30/dcn30\_dio\_stream\_encoder.c

\textbf{Function name 2:} enc3\_stream\_encoder\_stop\_hdmi\_info\_packets

\textbf{Similarity:} 50.33\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 8 }

\

\textbf{File 1:} dc/dml/dcn32/dcn32\_fpu.c

\textbf{Function name 1:} dcn32\_calculate\_wm\_and\_dlg\_fpu

\textbf{File 2:} dc/dml/dcn30/dcn30\_fpu.c

\textbf{Function name 2:} dcn30\_fpu\_update\_soc\_for\_wm\_a

\textbf{Similarity:} 50.81\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 9 }

\

\textbf{File 1:} dc/resource/dcn301/dcn301\_resource.c

\textbf{Function name 1:} dcn301\_dpp\_create

\textbf{File 2:} dc/resource/dcn321/dcn321\_resource.c

\textbf{Function name 2:} dcn321\_dpp\_destroy

\textbf{Similarity:} 50.39\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 10 }

\

\textbf{File 1:} dc/gpio/dce60/hw\_translate\_dce60.c

\textbf{Function name 1:} offset\_to\_id

\textbf{File 2:} dc/irq/dcn302/irq\_service\_dcn302.c

\textbf{Function name 2:} to\_dal\_irq\_source\_dcn302

\textbf{Similarity:} 49.33\%

\textbf{Duplication type:} Negative