### 2024-08-03 conteudo/apendices/noventa.tex

\section{Function pairs with similarity between 89\% and 91\%}

\paragraph{Function pair 1 }

\

\textbf{File 1:} dc/dce/dce\_opp.c

\textbf{Function name 1:} dce110\_opp\_program\_fmt

\textbf{File 2:} dc/dce/dce\_opp.c

\textbf{Function name 2:} dce60\_opp\_program\_fmt

\textbf{Similarity:} 89.59\%

\textbf{Duplication type:} Type-2

\paragraph{Function pair 2 }

\

\textbf{File 1:} dc/dpp/dcn10/dcn10\_dpp\_cm.c

\textbf{Function name 1:} dpp1\_cm\_get\_gamut\_remap

\textbf{File 2:} dc/dpp/dcn20/dcn20\_dpp\_cm.c

\textbf{Function name 2:} dpp2\_cm\_get\_gamut\_remap

\textbf{Similarity:} 90.20\%

\textbf{Duplication type:} Type-2

\paragraph{Function pair 3 }

\

\textbf{File 1:} dc/resource/dcn30/dcn30\_resource.c

\textbf{Function name 1:} dcn30\_hubbub\_create

\textbf{File 2:} dc/resource/dcn316/dcn316\_resource.c

\textbf{Function name 2:} dcn31\_hubbub\_create

\textbf{Similarity:} 89.02\%

\textbf{Duplication type:} Type-2

\paragraph{Function pair 4 }

\

\textbf{File 1:} dc/resource/dce110/dce110\_resource.c

\textbf{Function name 1:} dce110\_destroy\_resource\_pool

\textbf{File 2:} dc/resource/dce120/dce120\_resource.c

\textbf{Function name 2:} dce120\_destroy\_resource\_pool

\textbf{Similarity:} 90.39\%

\textbf{Duplication type:} Type-2

\paragraph{Function pair 5 }

\

\textbf{File 1:} dc/resource/dce120/dce120\_resource.c

\textbf{Function name 1:} dce120\_clock\_source\_create

\textbf{File 2:} dc/resource/dcn301/dcn301\_resource.c

\textbf{Function name 2:} dcn301\_clock\_source\_create

\textbf{Similarity:} 90.88\%

\textbf{Duplication type:} Type-2

\paragraph{Function pair 6 }

\

\textbf{File 1:} dc/dml/dcn32/dcn32\_fpu.c

\textbf{Function name 1:} dcn32\_enough\_pipes\_for\_subvp

\textbf{File 2:} dc/dml2/dml2\_mall\_phantom.c

\textbf{Function name 2:} enough\_pipes\_for\_subvp

\textbf{Similarity:} 90.74\%

\textbf{Duplication type:} Type-2

\paragraph{Function pair 7 }

\

\textbf{File 1:} dc/pg/dcn35/dcn35\_pg\_cntl.c

\textbf{Function name 1:} pg\_cntl35\_io\_clk\_status

\textbf{File 2:} dc/pg/dcn35/dcn35\_pg\_cntl.c

\textbf{Function name 2:} pg\_cntl35\_mem\_status

\textbf{Similarity:} 90.74\%

\textbf{Duplication type:} Type-2

\paragraph{Function pair 8 }

\

\textbf{File 1:} dc/hwss/dcn20/dcn20\_hwseq.c

\textbf{Function name 1:} dcn20\_disable\_writeback

\textbf{File 2:} dc/hwss/dcn30/dcn30\_hwseq.c

\textbf{Function name 2:} dcn30\_disable\_writeback

\textbf{Similarity:} 89.32\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 9 }

\

\textbf{File 1:} dc/dpp/dcn20/dcn20\_dpp\_cm.c

\textbf{Function name 1:} dpp20\_program\_shaper\_luta\_settings

\textbf{File 2:} dc/mpc/dcn32/dcn32\_mpc.c

\textbf{Function name 2:} mpc32\_program\_shaper\_lutb\_settings

\textbf{Similarity:} 89.34\%

\textbf{Duplication type:} Type-4

\paragraph{Function pair 10 }

\

\textbf{File 1:} dc/clk\_mgr/dcn31/dcn31\_smu.c

\textbf{Function name 1:} dcn31\_smu\_enable\_phy\_refclk\_pwrdwn

\textbf{File 2:} dc/clk\_mgr/dcn314/dcn314\_smu.c

\textbf{Function name 2:} dcn314\_smu\_enable\_phy\_refclk\_pwrdwn

\textbf{Similarity:} 89.03\%

\textbf{Duplication type:} Type-2