### 2024-08-03 conteudo/apendices/quarenta.tex

\section{Function pairs with similarity between 39\% and 41\%}

\paragraph{Function pair 1 }

\

\textbf{File 1:} dc/dc\_dmub\_srv.c

\textbf{Function name 1:} dc\_dmub\_srv\_is\_restore\_required

\textbf{File 2:} amdgpu\_dm/amdgpu\_dm.c

\textbf{Function name 2:} dm\_dmub\_hw\_resume

\textbf{Similarity:} 39.92\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 2 }

\

\textbf{File 1:} dc/hwss/dce110/dce110\_hwseq.c

\textbf{Function name 1:} dce110\_program\_front\_end\_for\_pipe

\textbf{File 2:} dc/hwss/dcn20/dcn20\_hwseq.c

\textbf{Function name 2:} dcn20\_enable\_stream

\textbf{Similarity:} 39.76\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 3 }

\

\textbf{File 1:} dc/core/dc\_resource.c

\textbf{Function name 1:} dc\_resource\_find\_first\_free\_pll

\textbf{File 2:} dc/core/dc\_resource.c

\textbf{Function name 2:} find\_free\_hpo\_dp\_link\_enc

\textbf{Similarity:} 40.64\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 4 }

\

\textbf{File 1:} dc/hwss/dcn10/dcn10\_hwseq.c

\textbf{Function name 1:} dcn10\_reset\_back\_end\_for\_pipe

\textbf{File 2:} dc/resource/dcn20/dcn20\_resource.c

\textbf{Function name 2:} dcn20\_add\_dsc\_to\_stream\_resource

\textbf{Similarity:} 39.00\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 5 }

\

\textbf{File 1:} dc/hubp/dcn20/dcn20\_hubp.c

\textbf{Function name 1:} hubp2\_read\_state\_common

\textbf{File 2:} dc/hubbub/dcn21/dcn21\_hubbub.c

\textbf{Function name 2:} hubbub21\_wm\_read\_state

\textbf{Similarity:} 40.60\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 6 }

\

\textbf{File 1:} dc/hwss/dcn20/dcn20\_hwseq.c

\textbf{Function name 1:} get\_phyd32clk\_src

\textbf{File 2:} dc/dml/calcs/dcn\_calcs.c

\textbf{Function name 2:} tl\_sw\_mode\_to\_bw\_defs

\textbf{Similarity:} 39.00\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 7 }

\

\textbf{File 1:} dc/hwss/dcn20/dcn20\_hwseq.c

\textbf{Function name 1:} dcn20\_init\_sys\_ctx

\textbf{File 2:} amdgpu\_dm/amdgpu\_dm.c

\textbf{Function name 2:} mmhub\_read\_system\_context

\textbf{Similarity:} 39.73\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 8 }

\

\textbf{File 1:} dc/clk\_mgr/dcn31/dcn31\_clk\_mgr.c

\textbf{Function name 1:} dcn31\_update\_clocks

\textbf{File 2:} dc/clk\_mgr/dcn30/dcn30\_clk\_mgr.c

\textbf{Function name 2:} dcn3\_init\_clocks

\textbf{Similarity:} 39.77\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 9 }

\

\textbf{File 1:} modules/power/power\_helpers.c

\textbf{Function name 1:} dmub\_init\_abm\_config

\textbf{File 2:} dc/hwss/dce110/dce110\_hwseq.c

\textbf{Function name 2:} dce110\_is\_hpo\_enabled

\textbf{Similarity:} 39.96\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 10 }

\

\textbf{File 1:} modules/hdcp/hdcp\_psp.c

\textbf{Function name 1:} mod\_hdcp\_add\_display\_to\_topology

\textbf{File 2:} modules/hdcp/hdcp\_ddc.c

\textbf{Function name 2:} mod\_hdcp\_read\_ksvlist

\textbf{Similarity:} 39.89\%

\textbf{Duplication type:} Negative