### 2024-08-03 conteudo/apendices/sesenta.tex

\section{Function pairs with similarity between 59\% and 61\%}

\paragraph{Function pair 1 }

\

\textbf{File 1:} dc/hwss/dcn30/dcn30\_hwseq.c

\textbf{Function name 1:} dcn30\_set\_avmute

\textbf{File 2:} dc/hwss/dcn10/dcn10\_hwseq.c

\textbf{Function name 2:} dcn10\_wait\_for\_pending\_cleared

\textbf{Similarity:} 59.99\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 2 }

\

\textbf{File 1:} dc/irq/dcn30/irq\_service\_dcn30.c

\textbf{Function name 1:} to\_dal\_irq\_source\_dcn30

\textbf{File 2:} dc/link/link\_detection.c

\textbf{Function name 2:} get\_basic\_signal\_type

\textbf{Similarity:} 60.97\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 3 }

\

\textbf{File 1:} dc/irq/dce60/irq\_service\_dce60.c

\textbf{Function name 1:} hpd\_ack

\textbf{File 2:} dc/irq/irq\_service.c

\textbf{Function name 2:} dal\_irq\_service\_set\_generic

\textbf{Similarity:} 60.92\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 4 }

\

\textbf{File 1:} dc/dml/dcn20/dcn20\_fpu.c

\textbf{Function name 1:} dcn20\_populate\_dml\_pipes\_from\_context

\textbf{File 2:} dc/resource/dcn314/dcn314\_resource.c

\textbf{Function name 2:} dcn314\_populate\_dml\_pipes\_from\_context

\textbf{Similarity:} 59.03\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 5 }

\

\textbf{File 1:} dc/dccg/dcn401/dcn401\_dccg.c

\textbf{Function name 1:} dccg401\_set\_physymclk

\textbf{File 2:} dc/dccg/dcn35/dcn35\_dccg.c

\textbf{Function name 2:} dccg35\_set\_physymclk

\textbf{Similarity:} 60.43\%

\textbf{Duplication type:} Type-4

\paragraph{Function pair 6 }

\

\textbf{File 1:} dc/hwss/dcn20/dcn20\_hwseq.c

\textbf{Function name 1:} dcn20\_enable\_plane

\textbf{File 2:} dc/hwss/dcn10/dcn10\_hwseq.c

\textbf{Function name 2:} dcn10\_enable\_plane

\textbf{Similarity:} 59.46\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 7 }

\

\textbf{File 1:} dc/dce/dce\_stream\_encoder.c

\textbf{Function name 1:} dce110\_stream\_encoder\_lvds\_set\_stream\_attribute

\textbf{File 2:} dc/dio/dcn10/dcn10\_stream\_encoder.c

\textbf{Function name 2:} enc1\_stream\_encoder\_dvi\_set\_stream\_attribute

\textbf{Similarity:} 60.67\%

\textbf{Duplication type:} Type-4

\paragraph{Function pair 8 }

\

\textbf{File 1:} dc/dml/dcn31/dcn31\_fpu.c

\textbf{Function name 1:} dcn31\_calculate\_wm\_and\_dlg\_fp

\textbf{File 2:} dc/dml/dcn20/dcn20\_fpu.c

\textbf{Function name 2:} dcn21\_calculate\_wm

\textbf{Similarity:} 59.46\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 9 }

\

\textbf{File 1:} dc/resource/dcn10/dcn10\_resource.c

\textbf{Function name 1:} dcn10\_resource\_destruct

\textbf{File 2:} dc/resource/dcn30/dcn30\_resource.c

\textbf{Function name 2:} dcn30\_resource\_construct

\textbf{Similarity:} 60.64\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 10 }

\

\textbf{File 1:} dc/clk\_mgr/dcn30/dcn30\_clk\_mgr.c

\textbf{Function name 1:} dcn3\_is\_smu\_present

\textbf{File 2:} dc/clk\_mgr/dcn316/dcn316\_clk\_mgr.c

\textbf{Function name 2:} dcn316\_enable\_pme\_wa

\textbf{Similarity:} 60.54\%

\textbf{Duplication type:} Negative