### 2024-08-03 conteudo/apendices/trinta.tex

\section{Function pairs with similarity between 29\% and 31\%}

\paragraph{Function pair 1 }

\

\textbf{File 1:} dc/core/dc\_state.c

\textbf{Function name 1:} dc\_state\_add\_all\_planes\_for\_stream

\textbf{File 2:} dc/resource/dcn20/dcn20\_resource.c

\textbf{Function name 2:} dcn20\_add\_dsc\_to\_stream\_resource

\textbf{Similarity:} 29.14\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 2 }

\

\textbf{File 1:} dc/gpio/dcn21/hw\_factory\_dcn21.c

\textbf{Function name 1:} define\_ddc\_registers

\textbf{File 2:} dc/link/protocols/link\_ddc.c

\textbf{Function name 2:} ddc\_service\_destruct

\textbf{Similarity:} 30.89\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 3 }

\

\textbf{File 1:} dc/core/dc\_resource.c

\textbf{Function name 1:} resource\_find\_free\_secondary\_pipe\_legacy

\textbf{File 2:} dc/dml/dcn32/dcn32\_fpu.c

\textbf{Function name 2:} update\_pipe\_slice\_table\_with\_split\_flags

\textbf{Similarity:} 29.61\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 4 }

\

\textbf{File 1:} dc/dml2/dml21/src/dml2\_core/dml2\_core\_dcn4.c

\textbf{Function name 1:} lookup\_uclk\_dpm\_index\_by\_freq

\textbf{File 2:} dc/dml2/dml2\_dc\_resource\_mgmt.c

\textbf{Function name 2:} is\_pipe\_in\_candidate\_array

\textbf{Similarity:} 30.65\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 5 }

\

\textbf{File 1:} dc/resource/dce112/dce112\_resource.c

\textbf{Function name 1:} dce112\_resource\_destruct

\textbf{File 2:} dc/resource/dcn20/dcn20\_resource.c

\textbf{Function name 2:} dcn20\_acquire\_dsc

\textbf{Similarity:} 30.99\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 6 }

\

\textbf{File 1:} modules/hdcp/hdcp\_ddc.c

\textbf{Function name 1:} mod\_hdcp\_write\_no\_stored\_km

\textbf{File 2:} modules/hdcp/hdcp\_ddc.c

\textbf{Function name 2:} mod\_hdcp\_write\_ainfo

\textbf{Similarity:} 29.82\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 7 }

\

\textbf{File 1:} dc/dml2/dml2\_mall\_phantom.c

\textbf{Function name 1:} all\_pipes\_have\_stream\_and\_plane

\textbf{File 2:} dc/basics/dce\_calcs.c

\textbf{Function name 2:} all\_displays\_in\_sync

\textbf{Similarity:} 30.59\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 8 }

\

\textbf{File 1:} dc/resource/dcn302/dcn302\_resource.c

\textbf{Function name 1:} dcn302\_resource\_destruct

\textbf{File 2:} dc/core/dc.c

\textbf{Function name 2:} dc\_resume

\textbf{Similarity:} 29.97\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 9 }

\

\textbf{File 1:} dc/optc/dcn314/dcn314\_optc.c

\textbf{Function name 1:} optc314\_enable\_crtc

\textbf{File 2:} dc/optc/dcn32/dcn32\_optc.c

\textbf{Function name 2:} optc32\_disable\_phantom\_otg

\textbf{Similarity:} 29.97\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 10 }

\

\textbf{File 1:} dc/irq/irq\_service.c

\textbf{Function name 1:} dal\_irq\_service\_ack

\textbf{File 2:} dc/irq/irq\_service.c

\textbf{Function name 2:} find\_irq\_source\_info

\textbf{Similarity:} 30.65\%

\textbf{Duplication type:} Negative