### 2024-11-14 conteudo/tool/tool.tex
add arkanjo as tool name

\en

Arkanjo, our proposed tool, is a Command Line Interface (CLI) application designed to
help developers identify code clone duplication at the function level.
Specifically, it enables the detection of pairs of functions within a codebase
that are clones of each other.

The tool operates in two main steps. The first step, called \textbf{Setup}, is
executed only once for each codebase. This step processes the codebase by
extracting functions from code files and identifying code clone duplications.
The second step, called \textbf{Query Responder}, utilizes the information obtained in
the Setup phase to execute the tool functionalities as needed. Detailed
descriptions of the Setup and Query Responder processes can be found in
Sections \ref{subsec:architecture} and \ref{subsec:func}.

The Arkanjo tool has been implemented and is available for reference at
\url{https://github.com/LipArcanjo/arkanjo}.