### 2024-11-15 conteudo/eval/table_empirical.tex
reforce that we do not expect good results in T3/T4 types

\begin{table}
\begin{tabular}{ | c | c | c | c | c | c | m{8em} | }

\hline

\textbf{Similarity range} & \textbf{T1} & \textbf{T2} & T3 & T4
& \textbf{Negative} & \textbf{Percentage} \\ \hline 
99\% - 100\% & 9 & 0 & 0 & 1 & 0 & \textbf{100\%} \\ \hline
89\% - 91\% & 0 & 8 & 0 & 1 & 1 & \textbf{90\%} \\ \hline
79\% - 81\% & 0 & 3 & 2 & 3 & 2 & \textbf{80\%} \\ \hline
69\% - 71\% & 0 & 3 & 1 & 1 & 5 & \textbf{50\%} \\ \hline
59\% - 61\% & 0 & 0 & 0 & 2 & 8 & \textbf{20\%} \\ \hline
49\% - 51\% & 0 & 0 & 0 & 0 & 0 & \textbf{0\%} \\ \hline
39\% - 41\% & 0 & 0 & 0 & 0 & 0 & \textbf{0\%} \\ \hline
29\% - 31\% & 0 & 0 & 0 & 0 & 10 & \textbf{0\%} \\ \hline

\hline

\end{tabular}
\caption{Results of the Arkanjo tool in the empirical analyses in the AMD Display driver}
\label{tab:emp}

The results were obtained by randomly selecting ten functions in the AMD Display driver for 
each similarity range and analyzing if they are duplications and which duplication type.

\end{table}