### 2025-05-03 conteudo/results/results.tex
reorginaze results files

\en

\todo[inline]{As it is primarly results, everything should be remade, the structure of the files for this section is bad.}

\todo[inline]{Remember the typos on the tables. }

\todo[inline]{Idea: The research methods should be 3 phases, one of the tool creation/testing, one for the my experiences submiting a path, and the last for the experiences of the students on the course. Maybe split in four to split the tool creation for the tool testing.}

To finalize this report, we present the current state of our research. As of the 
time of writing, we have finished \textbf{Phase I} of this research, 
which includes the implementation of 
the proposed tool and the evaluation of the tool through both a method from the 
literature and an empirical analyses. This chapter begins with the presentation of 
the results from the tool evaluation. Next, we present a preliminary experiment 
we are currently conducting to assess the viability of our Ethnographic Study. 
Finally, we outline the next steps in this research.

\input{conteudo/results/eval}

\input{conteudo/results/participant}