### 2025-05-11 conteudo/results/results.tex

\en

In this chapter we present and analyze the results found in the experiments and 
methods done through our research. Following the cronology of the research, we 
start by presenting the ArKanjo tool's evaluation through both a method from the literature
and an empirical analyses. Then, we present and discuss the experience of the initial participant
observation methodolody, where we mitigate a code duplication on the AMD Display driver and sent it as a patch.
Finally, we show and discuss our discoveries on the experiment done with students at the course in University of 
São Paulo, where we propose to students to mitigate duplications found by ArKanjo tool at IIO subsystem and the
AMD Display driver and document their founds, opinions and experiences.

\input{conteudo/results/eval}

\input{conteudo/results/participant}

\input{conteudo/results/students}