### 2025-05-11 conteudo/tool/tool.tex

\en

Arkanjo, our proposed tool, is a Command Line Interface (CLI) application designed to
help developers identify code clone duplication at the function level.
Specifically, it enables the detection of pairs of functions within a codebase
that are clones of each other.

The tool operates in two main parts, the \textbf{Preprocessor} and the \textbf{Query Responder}. 
The \textbf{Preprocessor} is responsible to perform the heavy computations to find the code duplications 
accross a codebase and produce artifact to consulte information related to the duplications in a structured form.
The \textbf{Query Responder} consumes the artifacts produced by the Preprocessor to execute the tool functionalites as requested.
This choice enable the tool to concentrate the heavy and slows parts that can be execute only once per codebase, 
thus allowing a fast performance for multiple queries related to a single codebase.
More detailed descriptions of the Preprocessor and Query Responder can be found in
Sections \ref{subsec:architecture} and \ref{subsec:func}.

The Arkanjo tool has been implemented and is available for reference at
\url{https://github.com/LipArcanjo/arkanjo}.

\todo[inline]{Big maybes: Describe more how the functionalites are implemented, as the focus shifted for the tool as of now.}