### 2025-05-25 conteudo/apendices/students-artifacts.tex
add some work from 24-05

\chapter{Students Artifacts}

\label{app:stuart}

This appendix includes the artifacts produced by the students at University of São Paulo.

\paragraph{Group 1}

\

\textbf{Blog 1:} \url{https://linux.ime.usp.br/~juliacalixto/software-livre.html}

\textbf{Blog 2:} \url{https://linux.ime.usp.br/~soueuabianca/post2.html}

\textbf{PATCH in Linux lore:} \url{https://lore.kernel.org/linux-iio/20250426194135.244774-1-biancalvao@gmail.com/}

\paragraph{Group 2}

\

\textbf{Blog 1:} \url{https://mikhael-abdallah.github.io/software-livre/posts/aula-06/}

\textbf{Blog 2:} \url{https://augustomb02.github.io/}

\textbf{PATCH in Linux lore:} \url{https://lore.kernel.org/all/20250429093923.347370-1-mikhael_abdallah@usp.br/}

\paragraph{Group 3}

\

\textbf{Blog 1:} \url{https://beavct.github.io/}

\textbf{Blog 2:} \url{https://gabriela-victor-github-io.vercel.app/blog/}

\textbf{PATCH in Linux lore:} \url{https://lore.kernel.org/linux-iio/20250424002144.23260-1-beatrizvianacosta16@gmail.com/}

\paragraph{Group 4}

\

\textbf{Blog 1:} \url{https://linux.ime.usp.br/~aschwarz}

\textbf{Blog 2:} \url{https://whip-foxtrot-b57.notion.site/MAC0470-Desenvolvimento-de-Software-Livre-2025-1dcde8837d}

\textbf{Blog 3:} \url{https://www.linux.ime.usp.br/~artpilone/sl_blog.html}

\textbf{PATCH in Linux lore:} \url{https://lore.kernel.org/linux-iio/20250421145534.91146-1-arthurpilone@usp.br/}

\paragraph{Group 5}

\

\textbf{Blog 1:} \url{https://gust-vaz.github.io/}

\textbf{Blog 2:} \url{https://xivor.github.io/}

\textbf{PATCH in Linux lore:} \url{https://lore.kernel.org/linux-iio/20250426153307.59689da0@jic23-huawei/T/#t}

\paragraph{Group 6}

\

\textbf{Blog 1:} \url{https://gabriellimmaa.github.io/}

\textbf{Blog 2:} \url{https://vitormarquesr.github.io/blog/}

\textbf{PATCH in Linux lore:} Students have not provided the link.

\paragraph{Group 7}

\

\textbf{Blog 1:} \url{https://isacaselli.github.io/}

\textbf{Blog 2:} \url{https://rodrigocmichelassi.github.io/}

\textbf{PATCH in Linux lore:} \url{https://lore.kernel.org/linux-iio/20250428132551.176788-1-bellacaselli20@gmail.com/}

\paragraph{Group 8}

\

\textbf{Blog 1:} \url{https://linux.ime.usp.br/~nattan/mac470/mac470.html}

\textbf{PATCH in Linux lore:} Students have not provided the link.

\paragraph{Group 9}

\

\textbf{Blog 1:} \url{https://andrewijano.github.io/mac5856/}

\textbf{PATCH in Linux lore:} \url{https://lore.kernel.org/linux-iio/20250423170526.15143-1-andrew.lopes@alumni.usp.br/}

\paragraph{Group 10}

\

\textbf{Blog 1:} \url{https://linux.ime.usp.br/~apolo/mac0470/}

\textbf{Blog 2:} \url{https://ncbor.github.io/}

\textbf{PATCH in Linux lore:} \url{https://lore.kernel.org/linux-iio/20250428034222.318898-1-gabrielgeraldinosouza@gmail.com/T/#u}

\paragraph{Group 11}

\

\textbf{Blog 1:} \url{https://o-carneiro.github.io/}

\textbf{Blog 2:} \url{https://fernandoglima.github.io/}

\textbf{Blog 3:} \url{https://eijiuchiyama.github.io/softwarelivre/}

\textbf{PATCH in Linux lore:} Students have not provided the link.

\paragraph{Group 12}

\

\textbf{Blog 1:} \url{https://saguileran.github.io/blog/2025/free-software-development/}

\textbf{PATCH in Linux lore:} \url{https://lore.kernel.org/amd-gfx/f78f68ba-80e0-47ca-97a2-37a15d076e68@amd.com/T/#t}