### 2025-06-01 conteudo/results/table_students.tex

\begin{table}
\begin{tabular}{ | c | c | c | c | m{6em} | }

\hline

\textbf{Group} & \textbf{Subsystem} & \textbf{Lines added} & \textbf{Lines removed} & \textbf{Refactoring methods used}
\\ \hline

1 & IIO & 7 & 14 & None \\ \hline
2 & IIO & 23 & 24 & Parameterize method \\ \hline
3 & IIO & 12 & 30 & Parameterize method \\ \hline
4 & IIO & 2 & 7 & None \\ \hline
5 & IIO & 17 & 70 & Parameterize method \\ \hline
6 & IIO & - & - & Parameterize method \\ \hline
7 & IIO & 21 & 22  & Extract method \\ \hline
8 & IIO & - & - & Extract method \\ \hline
9 & IIO & 33 & 52 & Parameterize method \\ \hline
10 & IIO & 32 & 58 & Parameterize method \\ \hline
11 & IIO & - & - & Parameterize method \\ \hline
12 & AMD & 89 & 489 & None \\ \hline

\hline
\end{tabular}
\caption{Refactor methods used and metrics of the students's mitigations.}
\label{tab:stu}
\end{table}