### 2025-06-01 conteudo/results/participant.tex

\en

\section{Participant Observation Experiment}

For our initial participant observation, we attempted to mitigate two code 
duplications found by the ArKanjo tool in the AMD Display driver by applying 
our systematic approach presented in \ref{subsubsec:systematic}. 
As presented in section \ref{sec:meteth},
we opted to apply the systematic approach in the functions
\textit{offset\_to\_id} and \textit{phy\_id\_to\_atom} in the folders 
\textit{dc/gpio} and \text{dc/bios}, respectively.

Hereafter, the mitigation on \textit{offset\_to\_id} will be referred to as mitigation 1, and the mitigation on 
\textit{phy\_id\_to\_atom} will be referred to as mitigation 2. 
Table \ref{tab:patch} summarizes the impact of the mitigations in terms of 
files and lines changed and the refactoring method used. In the next sections, 
we present our experience and learning for each mitigation.

\input{conteudo/results/table_patch}

\subsection{Mitigation 1}

In the systematic approach, we found that the \textit{offset\_to\_id} function 
exists in code 10 files, enclosing multiple GPU architectures. The duplicated 
functions are not exactly equal, as some minimal specific logic is applied to 
each GPU architecture. To address this issue, we used the extract method to 
break the function into smaller parts and the parameterized method to address 
the specific logic for the GPUs.

We reached a point with a scratch on how to mitigate the duplications found, 
but we could not complete the refactoring on a state capable of submitting as 
a patch to the driver. We found that the configuration files on the AMD Display 
driver are designed to be imported at compilation time using \textit{\#define} macros. 
The configuration files design choice makes the refactoring of duplications on 
generic approaches trickier, as refactoring code that depends on these files 
requires significant refactoring on the configuration files design and deep 
knowledge of the codebase, which we as first-timer contributors, do not have. 
Thus, we opted not to continue investigating this mitigation.

\subsection{Mitigation 2}

Using the systematic approach, we found that the \textit{phy\_id\_to\_atom}
function exists in five code files, and two other functions are duplicated 
in those five code files. All functions across the files were exactly equal, 
so there was no need to apply the refactoring methods presented in the literature. 
The refactoring work was resumed to create a generic library and fix the compilation targets.

The code files in the second mitigation do not depend on configuration files. Thus, 
we did not face the same issues during the first mitigation. Refactoring the code 
to mitigate duplications reached a point we judged was good enough to submit as a 
patch to contribute to the AMD Display driver code quality. Thus, we moved to send 
the refactoring to the driver’s maintainers as a patch, documenting our findings and experiences 
on the process.

\begin{figure}
\includegraphics[scale=0.45]{timeline_patch}
\caption{Timeline of the PATCH submission process}
\label{fig:timeline}
\end{figure}

The experience of submitting a patch was not as we expected. The process took 7 months 
and 16 days, which is significantly slower than expected. We initially submitted the 
patch on August 9th, 2024, but we did not receive replies. Thus, we had to resend it 
on October 9th, 2024. We received an initial response on November 3rd, 2024, asking for 
minimal changes; we submitted a second version with the changes request on November 16th, 2024. 
We did not receive replies at first and need to resend the patch on February 7th, 2025. 
We received a reply on February 11th, 2025, asking for additional changes. We sent a third 
version on February 24th, 2025, and the patch was accepted and integrated into the kernel 
codebase on February 25th, 2025. Figure \ref{fig:timeline} illustrates the timeline of 
the patch submission.

We discussed with our point of contact at the AMD Display driver to understand the reason 
for the slow process of submitting a patch. We learned that some of the code in the driver 
is shared across all operating systems that support the implemented GPUs. This fact creates 
a complex process within AMD to format and submit changes across the supported systems while 
simultaneously implementing measures to mitigate errors. Additionally, we comprehended that 
AMD Display driver developers can not view duplicated code negatively, as we understand it 
is a good practice in software engineering. One example shared with us is that duplicated 
code enhances the independence of GPU driver code, allowing developers to make changes to a 
specific GPU without needing to test compatibility with others. 
This approach helps save significant time and effort.

Regarding the patch changes requested for the second version, the maintainers only asked 
for minor changes to align with code style, license correct use, and best practices that 
were initially unknown to use. For the change requested on the third version, the maintainers 
asked us to move the functions created in the generic library to an existing file instead of 
creating a new one. More details of the patch can be found in Appendix \ref{app:feedback}, 
and the patch submitted in the Linux codebase can be seen here:
\url{https://lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br/}.

The initial plan of this research was to approach and send many mitigations to collect 
many artifacts from the driver community. Still, the slow process of sending a patch 
makes this plan unviable. Thus, we opted to redirect this research to work with students 
and refactoring duplicated function pairs found by the tool directly in the IIO and the 
AMD Display driver without passing through the whole process of the systematic approach 
proposed. This alternative allowed us to collect artifacts from the students' experiences 
and findings and not be too dependent on the interaction with the kernel community. 
Another advantage is accelerating the interaction with the kernel community because we are 
not applying the systematic approach, which translates to more minor code changes.