### 2025-06-12 conteudo/results/results.tex

\en

In this chapter, we present and analyze the results of the experiments and methods 
used in our research. Following the chronology of the study, we start by presenting the 
evaluation of the ArKanjo tool through both a method from the literature and an empirical 
analysis. Then, we present and discuss the experience of the initial participant 
observation methodology, where we mitigate a code duplication on the AMD Display Driver 
and send it as a patch. Finally, we show and discuss our discoveries of the non-participant observation
of the students approaches to mitigate duplications and patches's discussions.

\input{conteudo/results/eval}

\input{conteudo/results/participant}

\input{conteudo/results/students}