### 2025-06-12 conteudo/results/table_students.tex

\begin{table}
\begin{tabular}{ | c | c | c | m{6em} | }

\hline

\textbf{Group} & \textbf{Subsystem} & \textbf{Patch Status} & \textbf{Refactoring methods used} 
\\ \hline

1 & IIO & Rejected & None \\ \hline
2 & IIO & Rejected & Parameterize method \\ \hline
3 & IIO & Accepted & Parameterize method \\ \hline
4 & IIO & Accepted & None \\ \hline
5 & IIO & Accepted  & Parameterize method \\ \hline
6 & IIO & Students stopped working on it & Parameterize method \\ \hline
7 & IIO & Accepted  & Extract method \\ \hline
8 & IIO & Rejected & Extract method \\ \hline
9 & IIO & Work in progress & Parameterize method \\ \hline
10 & IIO & Rejected & Parameterize method \\ \hline
11 & IIO & Students stopped working on it & Parameterize method \\ \hline
12 & AMD & Accepted & None \\ \hline

\hline
\end{tabular}
\caption{Refactor methods used and metrics of the students's mitigations.}
\label{tab:stu}
\end{table}