### 2025-08-05 conteudo/conclusion/conclusion.tex
Review conclusion

The Linux Kernel massive scale, with over 28 million lines of code 
and more than twenty thousand contributors, presents significant maintenance 
challenges; code duplication persists as a notable issue within its codebase, 
potentially hindering its evolution and patching processes. Prevailing academic 
and existing FLOSS approaches to duplication detection often lack the 
scalability for comprehensive codebase analysis or offer limited functionality 
unsuitable for the kernel complexity, generally failing to deliver actionable 
results or guidance for mitigation within its specific context.

In response to these limitations, this research addressed the challenge of code 
duplication within the Linux Kernel by developing ArKanjo, a command-line tool 
for Linux kernel maintenance designed to detect and analyze function-level duplications. 
Released under the MIT license, the architectural design of ArKanjo, incorporating a 
Preprocessor and a Query Responder, separates computationally intensive analysis from 
efficient querying for duplications within large codebases. Validation procedures, 
employing both the BigCloneBench dataset and empirical analysis conducted on the AMD 
Display driver, indicated ArKanjo's proficient performance in identifying Type-1 and 
Type-2 duplications, notably when similarity thresholds are configured between 80\% and 100\%.

The ethnographic studies, which involved participant observation and an experiment with university 
students, provided a realistic view of the opportunities and the considerable challenges in 
mitigating code duplications found by ArKanjo. The tool was effectively used by first-time 
contributors to the Linux kernel to 
identify duplications and prepare patches. While this research included a successfully merged 
patch by the author, and successful submissions from many student groups, a notable portion of 
student efforts encountered significant hurdles. These difficulties often stemmed from maintainer 
feedback where proposed changes, though reducing duplication, were rejected or required 
substantial rework due to concerns about code readability, added abstraction, or the specific 
context of the code. Technical complexities, such as C macros in configuration files, and lengthy 
patch review times also contributed to the challenges. Furthermore, the studies confirmed that 
maintainers do not always view code duplication as inherently harmful, especially when it may 
improve clarity or facilitate independent hardware support.

This research contributes to the ArKanjo tool as an operational method for detecting code 
duplication and, importantly, provides insights into the process of managing such duplications 
within the Linux kernel subsystems studied. The results show that while ArKanjo can effectively 
identify duplicated functions, the path from detection to an accepted mitigation within the kernel 
involves careful consideration of the specific code, established community practices, and the 
maintainers' perspectives on code quality and readability.

\section{Future Work}

The work presented in this thesis suggests several directions for future investigation. Enhancements 
to the ArKanjo tool could be pursued, such as investigating a multi-stage detection strategy. 
This strategy might employ a computationally lighter method for an initial, broad pass to identify 
candidate duplications, followed by more resource-intensive, state-of-the-art techniques for refined 
analysis and improved detection of complex clone types. Such an approach could complement efforts for 
improved resource management during its preprocessing phase for large codebases. The tool language 
support could be broadened by developing parsers for additional programming languages, building upon 
its adaptable architecture. Beyond the tool itself, future research could extend the empirical analysis 
of code duplication and mitigation strategies to a broader range of Linux kernel subsystems or other 
large-scale C projects, to determine the broader applicability of the observations made.