### 2025-10-28 masterthesis/conteudo/apendices/emp-lista.tex

\chapter{Random selected samples}

\label{app:emp}

This appendix includes the list of function pairs used to evaluate the tool, as 
described in section \ref{sec:metemp}. 
The results are available in Table \ref{tab:emp}. 
The files indicate the relative path to the AMD Display driver root directory.

\input{conteudo/apendices/tool/cem}

\input{conteudo/apendices/tool/noventa}

\input{conteudo/apendices/tool/oitenta}

\input{conteudo/apendices/tool/setenta}

\input{conteudo/apendices/tool/sesenta}

\input{conteudo/apendices/tool/cinquenta}

\input{conteudo/apendices/tool/quarenta}

\input{conteudo/apendices/tool/trinta}