### 2025-10-28 masterthesis/conteudo/apendices/students-artifacts.tex

\chapter{Students Artifacts}

\label{app:stuart}

This appendix includes the artifacts produced by the students at University of São Paulo.

\paragraph{Group 1}

\

\textbf{Blog 1:} \textit{linux.ime.usp.br/~juliacalixto/software-livre.html}

\textbf{Blog 2:} \textit{linux.ime.usp.br/~soueuabianca/post2.html}

\textbf{PATCH in Linux lore:} \textit{\seqsplit{lore.kernel.org/linux-iio/20250426194135.244774-1-biancalvao@gmail.com/}}

\paragraph{Group 2}

\

\textbf{Blog 1:} \textit{mikhael-abdallah.github.io/software-livre/posts/aula-06/}

\textbf{Blog 2:} \textit{augustomb02.github.io/}

\textbf{PATCH in Linux lore:} \textit{\seqsplit{lore.kernel.org/all/20250429093923.347370-1-mikhael\_abdallah@usp.br/}}

\paragraph{Group 3}

\

\textbf{Blog 1:} \textit{beavct.github.io/}

\textbf{Blog 2:} \textit{gabriela-victor-github-io.vercel.app/blog/}

\textbf{PATCH in Linux lore:} \textit{\seqsplit{lore.kernel.org/linux-iio/20250424002144.23260-1-beatrizvianacosta16@gmail.com/}}

\textbf{Accepted PATCH (v1):} \textit{\seqsplit{lore.kernel.org/linux-iio/20250424002144.23260-1-beatrizvianacosta16@gmail.com/}}

\paragraph{Group 4}

\

\textbf{Blog 1:} \textit{linux.ime.usp.br/~aschwarz}

\textbf{Blog 2:} \textit{\seqsplit{whip-foxtrot-b57.notion.site/MAC0470-Desenvolvimento-de-Software-Livre-2025-1dcde8837d}}

\textbf{Blog 3:} \textit{www.linux.ime.usp.br/~artpilone/sl\_blog.html}

\textbf{PATCH in Linux lore:} \textit{\seqsplit{lore.kernel.org/linux-iio/20250421145534.91146-1-arthurpilone@usp.br/}}

\textbf{Accepted PATCH (v3):} \textit{\seqsplit{lore.kernel.org/linux-iio/20250426163311.546500bc@jic23-huawei/}}

\paragraph{Group 5}

\

\textbf{Blog 1:} \textit{gust-vaz.github.io/}

\textbf{Blog 2:} \textit{xivor.github.io/}

\textbf{PATCH in Linux lore:} \textit{\seqsplit{lore.kernel.org/linux-iio/20250426153307.59689da0@jic23-huawei/T/\#t}}

\textbf{Accepted PATCH (v1):} \textit{\seqsplit{lore.kernel.org/linux-iio/20250426153307.59689da0@jic23-huawei/T/\#t}}

\paragraph{Group 6}

\

\textbf{Blog 1:} \textit{gabriellimmaa.github.io/}

\textbf{Blog 2:} \textit{vitormarquesr.github.io/blog/}

\textbf{PATCH in Linux lore:} \textit{\seqsplit{lore.kernel.org/all/20250625101857.00004dbf@huawei.com/}}

\paragraph{Group 7}

\

\textbf{Blog 1:} \textit{isacaselli.github.io/}

\textbf{Blog 2:} \textit{rodrigocmichelassi.github.io/}

\textbf{PATCH in Linux lore:} \textit{\seqsplit{lore.kernel.org/linux-iio/20250428132551.176788-1-bellacaselli20@gmail.com/}}

\textbf{Accepted PATCH (v3):} \textit{\seqsplit{lore.kernel.org/linux-iio/20250531160807.39e54096@jic23-huawei/}}

\paragraph{Group 8}

\

\textbf{Blog 1:} \textit{linux.ime.usp.br/~nattan/mac470/mac470.html}

\textbf{PATCH in Linux lore:} \textit{\seqsplit{lore.kernel.org/all/20250422191114.32832-1-nattanferreira58@gmail.com/\#t}}

\textbf{Accepted PATCH (v4):} \textit{\seqsplit{lore.kernel.org/all/20250614124401.32b17eda@jic23-huawei/}}

\paragraph{Group 9}

\

\textbf{Blog 1:} \textit{andrewijano.github.io/mac5856/}

\textbf{PATCH in Linux lore:} \textit{\seqsplit{lore.kernel.org/linux-iio/20250423170526.15143-1-andrew.lopes@alumni.usp.br/}}

\textbf{Accepted PATCH (partially accepted on v6):} \textit{\seqsplit{lore.kernel.org/linux-iio/20250621183843.2f8bcb48@jic23-huawei/}}

\paragraph{Group 10}

\

\textbf{Blog 1:} \textit{linux.ime.usp.br/~apolo/mac0470/}

\textbf{Blog 2:} \textit{ncbor.github.io/}

\textbf{PATCH in Linux lore:} \textit{\seqsplit{lore.kernel.org/linux-iio/20250428034222.318898-1-gabrielgeraldinosouza@gmail.com/T/\#u}}

\paragraph{Group 11}

\

\textbf{Blog 1:} \textit{\seqsplit{saguileran.github.io/blog/2025/free-software-development/}}

\textbf{PATCH in Linux lore:} \textit{\seqsplit{lore.kernel.org/amd-gfx/f78f68ba-80e0-47ca-97a2-37a15d076e68@amd.com/T/\#t}}

\textbf{Accepted PATCH (v2):} \textit{\seqsplit{https://lore.kernel.org/amd-gfx/CADAW+GDN-4Lwd87h793csWF6r5cgqpgrpEPzRyDV9+rfkNPmTA@mail.gmail.com/}}