### 2025-10-28 masterthesis/conteudo/apendices/tool/oitenta.tex

\section{Function pairs with similarity between 79\% and 81\%}

\paragraph{Function pair 1 }

\

\textbf{File 1:} dc/dpp/dcn401/dcn401\_dpp.c

\textbf{Function name 1:} dpp401\_dpp\_setup

\textbf{File 2:} dc/dpp/dcn10/dcn10\_dpp.c

\textbf{Function name 2:} dpp1\_cnv\_setup

\textbf{Similarity:} 79.60\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 2 }

\

\textbf{File 1:} dc/dml/dcn32/dcn32\_fpu.c

\textbf{Function name 1:} dcn32\_calculate\_wm\_and\_dlg\_fpu

\textbf{File 2:} dc/dml/dcn31/dcn31\_fpu.c

\textbf{Function name 2:} dcn31\_calculate\_wm\_and\_dlg\_fp

\textbf{Similarity:} 80.41\%

\textbf{Duplication type:} Type-4

\paragraph{Function pair 3 }

\

\textbf{File 1:} dc/resource/dcn316/dcn316\_resource.c

\textbf{Function name 1:} dcn316\_resource\_construct

\textbf{File 2:} dc/resource/dce80/dce80\_resource.c

\textbf{Function name 2:} dce81\_construct

\textbf{Similarity:} 79.37\%

\textbf{Duplication type:} Type-4

\paragraph{Function pair 4 }

\

\textbf{File 1:} dc/mpc/dcn401/dcn401\_mpc.c

\textbf{Function name 1:} get3dlut\_config

\textbf{File 2:} dc/dpp/dcn20/dcn20\_dpp\_cm.c

\textbf{Function name 2:} get3dlut\_config

\textbf{Similarity:} 80.62\%

\textbf{Duplication type:} Type-2

\paragraph{Function pair 5 }

\

\textbf{File 1:} dc/resource/dce100/dce100\_resource.c

\textbf{Function name 1:} dce100\_hwseq\_create

\textbf{File 2:} dc/resource/dce120/dce120\_resource.c

\textbf{Function name 2:} dce121\_hwseq\_create

\textbf{Similarity:} 79.34\%

\textbf{Duplication type:} Type-2

\paragraph{Function pair 6 }

\

\textbf{File 1:} dc/resource/dce80/dce80\_resource.c

\textbf{Function name 1:} dce80\_construct

\textbf{File 2:} dc/resource/dcn314/dcn314\_resource.c

\textbf{Function name 2:} dcn314\_resource\_construct

\textbf{Similarity:} 79.19\%

\textbf{Duplication type:} Type-4

\paragraph{Function pair 7 }

\

\textbf{File 1:} dc/clk\_mgr/dcn315/dcn315\_clk\_mgr.c

\textbf{Function name 1:} dcn315\_clk\_mgr\_destroy

\textbf{File 2:} dc/clk\_mgr/dcn35/dcn35\_clk\_mgr.c

\textbf{Function name 2:} dcn35\_clk\_mgr\_destroy

\textbf{Similarity:} 79.16\%

\textbf{Duplication type:} Type-2

\paragraph{Function pair 8 }

\

\textbf{File 1:} dc/clk\_mgr/dcn401/dcn401\_clk\_mgr\_smu\_msg.c

\textbf{Function name 1:} dcn401\_smu\_send\_msg\_with\_param

\textbf{File 2:} dc/clk\_mgr/dcn30/dcn30\_clk\_mgr\_smu\_msg.c

\textbf{Function name 2:} dcn30\_smu\_send\_msg\_with\_param

\textbf{Similarity:} 80.30\%

\textbf{Duplication type:} Type-3

\paragraph{Function pair 9 }

\

\textbf{File 1:} dc/link/link\_dpms.c

\textbf{Function name 1:} link\_set\_dpms\_off

\textbf{File 2:} dc/link/link\_dpms.c

\textbf{Function name 2:} link\_set\_dpms\_on

\textbf{Similarity:} 80.74\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 10 }

\

\textbf{File 1:} dc/clk\_mgr/dcn30/dcn30\_clk\_mgr\_smu\_msg.c

\textbf{Function name 1:} dcn30\_smu\_set\_hard\_min\_by\_freq

\textbf{File 2:} dc/clk\_mgr/dcn401/dcn401\_clk\_mgr\_smu\_msg.c

\textbf{Function name 2:} dcn401\_smu\_set\_hard\_min\_by\_freq

\textbf{Similarity:} 80.97\%

\textbf{Duplication type:} Type-3