### 2025-10-28 masterthesis/conteudo/apendices/tool/setenta.tex

\section{Function pairs with similarity between 69\% and 71\%}

\paragraph{Function pair 1 }

\

\textbf{File 1:} dc/resource/dcn21/dcn21\_resource.c

\textbf{Function name 1:} read\_dce\_straps

\textbf{File 2:} dc/resource/dcn321/dcn321\_resource.c

\textbf{Function name 2:} read\_dce\_straps

\textbf{Similarity:} 69.52\%

\textbf{Duplication type:} Type-2

\paragraph{Function pair 2 }

\

\textbf{File 1:} amdgpu\_dm/amdgpu\_dm\_pp\_smu.c

\textbf{Function name 1:} pp\_rv\_set\_min\_deep\_sleep\_dcfclk

\textbf{File 2:} amdgpu\_dm/amdgpu\_dm\_pp\_smu.c

\textbf{Function name 2:} pp\_rv\_set\_hard\_min\_fclk\_by\_freq

\textbf{Similarity:} 69.11\%

\textbf{Duplication type:} Type-2

\paragraph{Function pair 3 }

\

\textbf{File 1:} dc/hubbub/dcn35/dcn35\_hubbub.c

\textbf{Function name 1:} convert\_and\_clamp

\textbf{File 2:} dc/dml/dml1\_display\_rq\_dlg\_calc.c

\textbf{Function name 2:} is\_dual\_plane

\textbf{Similarity:} 69.41\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 4 }

\

\textbf{File 1:} dc/irq/dcn35/irq\_service\_dcn35.c

\textbf{Function name 1:} to\_dal\_irq\_source\_dcn35

\textbf{File 2:} dc/dml/calcs/dcn\_calcs.c

\textbf{Function name 2:} tl\_sw\_mode\_to\_bw\_defs

\textbf{Similarity:} 70.65\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 5 }

\

\textbf{File 1:} dc/spl/dc\_spl\_scl\_filters.c

\textbf{Function name 1:} spl\_get\_filter\_4tap\_16p

\textbf{File 2:} dc/spl/dc\_spl\_scl\_filters.c

\textbf{Function name 2:} spl\_get\_filter\_5tap\_64p

\textbf{Similarity:} 70.47\%

\textbf{Duplication type:} Type-2

\paragraph{Function pair 6 }

\

\textbf{File 1:} include/fixed31\_32.h

\textbf{Function name 1:} dc\_fixpt\_max

\textbf{File 2:} dc/dml/calcs/dcn\_calc\_math.c

\textbf{Function name 2:} dcn\_bw\_max

\textbf{Similarity:} 70.50\%

\textbf{Duplication type:} Type-3

\paragraph{Function pair 7 }

\

\textbf{File 1:} amdgpu\_dm/amdgpu\_dm\_crc.c

\textbf{Function name 1:} dm\_need\_crc\_dither

\textbf{File 2:} amdgpu\_dm/amdgpu\_dm\_crc.c

\textbf{Function name 2:} dm\_is\_crc\_source\_dprx

\textbf{Similarity:} 69.73\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 8 }

\

\textbf{File 1:} dc/optc/dcn35/dcn35\_optc.c

\textbf{Function name 1:} optc35\_set\_drr

\textbf{File 2:} dc/optc/dcn35/dcn35\_optc.c

\textbf{Function name 2:} optc35\_set\_long\_vtotal

\textbf{Similarity:} 70.46\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 9 }

\

\textbf{File 1:} dc/resource/dcn201/dcn201\_resource.c

\textbf{Function name 1:} dcn201\_hwseq\_create

\textbf{File 2:} dc/resource/dcn10/dcn10\_resource.c

\textbf{Function name 2:} dcn10\_hwseq\_create

\textbf{Similarity:} 69.21\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 10 }

\

\textbf{File 1:} dc/resource/dce80/dce80\_resource.c

\textbf{Function name 1:} dce80\_link\_encoder\_create

\textbf{File 2:} dc/resource/dcn10/dcn10\_resource.c

\textbf{Function name 2:} dcn10\_link\_encoder\_create

\textbf{Similarity:} 70.72\%

\textbf{Duplication type:} Type-4