### 2025-10-28 masterthesis/conteudo/results/results.tex

\en

In this chapter, we present and analyze the results of the experiments and methods 
used in our research. Following the chronology of the study, we start by presenting 
the evaluation of the ArKanjo tool through both a method from the literature and an empirical
analysis. Then, we present and discuss the experience of the initial participant 
observation methodology, where we mitigate one code duplication on the AMD Display 
driver and send it as a patch. Finally, we show and discuss our discoveries of 
the non-participant observation of the students' approaches to mitigate duplications 
and patches' discussions.

\input{conteudo/results/eval}

\input{conteudo/results/participant}

\input{conteudo/results/students}