### 2025-10-28 masterthesis/conteudo/tool/tool.tex

\en

ArKanjo, our proposed tool, is a Command Line Interface (CLI) application designed to
help developers identify code clone duplication at the function level.
Specifically, it enables the detection of pairs of functions that are clones of 
each other within a codebase.

The tool operates in two main parts: the \textbf{Preprocessor} and the \textbf{Query Responder}. 
The \textbf{Preprocessor} performs heavy computations to find the code 
duplications across a codebase and produce artifacts to consult information related to the 
duplications in a structured form. The \textbf{Query Responder} consumes the artifacts produced 
by the Preprocessor to execute the tool functionalities as requested. This choice enables the 
tool to concentrate on the heavy and slow parts that can be executed only once per codebase, 
thus allowing a fast performance for multiple queries related to a single codebase.

The ArKanjo tool has been implemented and is available for reference at
\url{github.com/arkanjo-tool/arkanjo}.