### 2025-10-29 papers/techdebt2026/techdebt.tex
techdebt2026: techdebt: revise results section
Signed-off-by: David Tadokoro <davidbtadokoro@usp.br>

\documentclass[10pt,conference]{IEEEtran}
\usepackage{graphicx,url}
\usepackage{hyperref}
\usepackage{nicematrix}
\usepackage{array}
\usepackage[utf8]{inputenc}  
\usepackage{seqsplit}
\usepackage{todonotes}

%% Takeaway Boxes %%
\usepackage[most]{tcolorbox}
\tcbset{colback=gray!10, colframe=gray!50!black, boxrule=0.5pt,
        left=1mm, right=1mm, top=1mm, bottom=1mm, sharp corners}

\newenvironment{highlight-box}[1]{%
  \begin{tcolorbox}
  \textbf{#1:} \itshape}{\end{tcolorbox}}
%% Takeaway Boxes %%

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}

\title{When Do You Repeat Yourself?\\
Voices from the Trenches of Linux Kernel Maintainers on Code Duplication}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations

\author{\IEEEauthorblockN{Anonymous Authors}}

%\author{\IEEEauthorblockN{Luan Arcanjo, Marcelo Spessoto, David Tadokoro, Paulo Meirelles}
%\IEEEauthorblockA{
%Free Software Competence Center\\
%Institute of Mathematics, Statistics, and Computer Science
%Universidade de São Paulo, Brazil\\
%\{luanicaro,marcelomspessoto,davidbtadokoro\}@usp.br,paulormm@ime.usp.br}
%}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
The vast scale and continuous evolution of the Linux kernel make its maintenance a complex undertaking, where code duplication remains a persistent challenge that can hinder development and introduce bugs.
%
This paper presents a multimethod ethnographic study investigating the socio-technical dynamics of contributing to the Linux kernel by reducing identified code duplication.
%
To support this study, we developed a command-line utility tool for detecting function-level duplications, offering a concrete entry point for new contributors.
%
Our ethnographic approach, including complete participant observation and participant-as-observer analyses, demonstrated that addressing duplications is viable for lowering the contribution barrier. This point is evidenced by 8 of 13 (62\%) accepted contributions (patches) from 24 newcomers (16 undergraduate and 8 graduate students) in the Linux kernel project, collectively removing 1.397 lines of duplicated code.
%
Nevertheless, the study reveals a more intriguing and complex reality beyond mindlessly removing clones. An analysis of maintainer feedback on both accepted and rejected contributions highlights a nuanced understanding of technical debt due to code duplication within the Linux kernel community, where the benefits of eliminating duplications are carefully weighed against factors such as readability, the introduction of new abstractions, and the specific code context.
\end{abstract}

% no keywords

\IEEEpeerreviewmaketitle

\section{Introduction}
\label{sec:introduction}

Hundreds or thousands of developers working together on a software product is a complex and demanding task. Similar or even exact copies are a frequent and often unavoidable outcome. Excessive duplication can negatively impact the maintainability of a project, as, when it occurs, any change to a repeated code segment (to fix a bug, for instance) requires replication across all other copies. Developers must, therefore, remain constantly vigilant to synchronize different parts of the codebase, which is a highly error-prone yet seemingly simple responsibility. Regarded as a major source of technical debt, removing or refactoring duplicated code demands great care to preserve the original behavior, with minor mistakes possibly leading to serious problems in the future.

A widespread dogma among developers is the \textit{Don't Repeat Yourself} principle~\cite{pragmatic-programmer}, stated as follows.

\begin{highlight-box}{DRY principle}
  ``We feel that the only way to develop software reliably, and to make our
  developments easier to understand and maintain, is to follow what we call the
  DRY principle: Every piece of knowledge must have a single, unambiguous,
  authoritative representation within a system''.
\end{highlight-box}

The Linux kernel is a foundational Free/Libre and Open Source Software (FLOSS) project, a vital component of the world's digital infrastructure. Considering only lines of code (no comments or blank lines) of strictly programming languages, in version 6.17, the main repository has more than 27 million lines of code and involved contributions from over twenty thousand developers. Within this context, code duplication remains a significant challenge,  reducing the code readability and increasing the risk of introducing bugs~\cite{harmone,harmtwo}. This issue is particularly prominent in kernel device drivers, which account for over 66\% of the Linux kernel source code.

The detection of code duplication, or \textit{code clones}, has been a research topic for decades~\cite{firstman}. The literature offers a widely accepted taxonomy that classifies clones into four types based on their similarity, from exact copies (Type-1) to semantically equivalent but syntactically different fragments (Type-4)~\cite{litreview}. Several detection techniques have emerged, including textual, token-based, tree-based, and graph-based approaches~\cite{litreview}. These efforts have culminated in state-of-the-art techniques, such as the graph-based approach proposed by Liu et al.~\cite{tailor}.

To address code duplication in the Linux kernel, we developed a command-line tool capable of detecting and analyzing function-level duplications\footnote{The tool and its documentation will be referenced after the review process.}. We conducted a multimethod ethnographic study leveraging the tool capabilities to identify duplications and submit contributions, directly or indirectly, to the Linux kernel, aiming to remove redundant code.
%
Through interactions with maintainers, this study gained insights into how the Linux community perceives the code quality related to code clones. Contrary to the DRY principle, our findings indicate that, in some contexts, maintainers can tolerate duplications for better readability, performance increases, or even to avoid integration overhead.

Ethnography is a well-established qualitative method for understanding people, their cultures, and work practices~\cite{bookethno}. It provides insights into community members' values, beliefs, and practices~\cite{ethnosoft}. In this study, we first employed a complete participant observation approach~\cite{gold1958roles}, actively contributing by performing deduplications while documenting the process. We then conducted a participant-as-observer study~\cite{gold1958roles} with students resolving duplications identified by the tool. Beyond demonstrating that the tool lowered the entry barrier for new contributors, these studies yielded intriguing observations on kernel Linux deduplications. This paper describes the aforementioned studies, along with our visions and reflections from diving into the Linux kernel development process.

\section{The Linux Kernel}

The Linux kernel is a complex project organized into subsystems, such as the process scheduler, memory management, and device drivers. Each subsystem typically has a dedicated maintainer or a team of maintainers overseeing its development and managing contributions. The development process is coordinated through Git (also designed by Linus Torvald, focused on the Linux kernel development), and contributions are formatted as patches, text documents that outline the differences between two source code versions. These patches are then submitted to the relevant Linux mailing lists for public review and discussion.

This study focuses on two of these subsystems, selected to represent different contexts for contribution. We explore the \textit{AMD Display drivers}, chosen for their importance within the hardware ecosystem, which is responsible for natively enabling AMD GPU functionality in Linux. This subsystem is substantial, with over 391,000 lines of code across more than 1,089 files. We also investigate the \textit{Industrial Input/Output (IIO)} subsystem, often considered an accessible entry point for new contributors. The IIO subsystem provides support for devices that perform analog-to-digital or digital-to-analog conversions\footnote{\href{https://kernel.org/doc/html/v6.17/driver-api/iio/intro.html}{https://kernel.org/doc/html/v6.17/driver-api/iio/intro.html}} and contains over 281,772 lines of code in more than 755 files.

\section{Research Strategy}

We employed a multi-method research approach. We developed and evaluated a proposed tool using a method from the literature~\cite{bigclonebench}. Given this validation, we proceeded to understand the viability of reducing duplications found by the tool with an ethnographic study using a participant observation methodology~\cite{gold1958roles}. In the first phase, we interacted with the Linux kernel community and got their feedback (\textit{complete participant observation}~\cite{gold1958roles}) through sending deduplication patches to the AMD Display driver. In the second phase, during a university course, we proposed a similar activity of sending patches to AMD Display or IIO to students. At the same time, we guided and mentored them throughout the process (\textit{participant-as-observer}~\cite{gold1958roles}).

\subsection{Tool}

Academic research on code duplication often focuses on pairwise comparison of code artifacts, which is not directly applicable to comprehensive codebase analyses. Existing free software tools we explored in practice suffered from limited functionality~[omitted]. Thus, we developed a command-line tool\footnote{Anonymous repository:  \url{https://anonymous.4open.science/r/arkanjo-C548}} designed to detect and analyze function-level code duplication within large codebases, such as the Linux kernel. The tool is released under LGPLv3 (GNU Lesser General Public License, version 3) and employs a two-stage architecture that separates the computationally intensive task of finding duplicates from the process of querying the results. The first stage analyzes the entire codebase, parses C functions, and stores identified duplicates in a database. The second stage uses this database to retrieve information about the identified duplications quickly.
%The tool is available at \textit{\href{https://github.com/arkanjo-tool/arkanjo}{github.com/arkanjo-tool/arkanjo}}.

The tool was designed to address the specific challenges posed by large-scale codebases, focusing on scalability and prioritization of duplications that may impact maintainability. It uses a text similarity method based on TF-IDF (Term Frequency-Inverse Document Frequency) vector embedding, implemented with the Gensim library~\cite{gensim}, and computes cosine similarity to detect clones.

We compared our tool against the BigCloneBench dataset~\cite{bigclonebench}. This validation showed that the tool is capable of detecting the majority of Type-1 and Type-2 duplications (100\% of Type-1 and 85\% of Type-2 with a cosine similarity threshold of 0.9), demonstrating sufficient accuracy to support our ethnographic studies.

\subsection{Complete Participant Observation Study}
\label{sec:participant}

To investigate whether we could use the tool to find ways to contribute to the Linux kernel, we first conducted a complete participant observation experiment, where the main author (a graduate student) acted as a first-time contributor to the AMD Display driver, collecting artifacts of our experience in the process.

We executed the tool in the AMD Display driver codebase. We manually analyzed the biggest duplications regarding the number of lines, finding one pair that we judged promising to try to deduplicate. We approached big duplications as proof of concept to gauge the tool's potential for enabling significant impacts.

Given the duplicated function pair, we observed that the files in the functions contained multiple other duplications. Thus, we proposed a simple systematic approach to mitigate all the functions duplicated in the context, not just the duplicated function pairs. After approaching the duplications with a systematic strategy, we sent a patch to the AMD development mailing list for the feedback of maintainers, while documenting the process, interactions, and impressions (available in our Replication Package).

In C programming, communication between source files is achieved by creating header files that specify libraries~\cite{Cbook}. Since the Linux kernel is primarily written in C, our systematic strategy consisted of eliminating code duplications through consolidating the duplicated code into a single library, which would replace instances of duplication across the codebase. For each duplicated function pair, for each function in the pair, we used the tool to locate all duplications of the function in the codebase, as it could have been duplicated more times than the two occurrences detected initially. This strategy resulted in a collection of duplicated functions and corresponding code files.

To identify shared code more effectively, we extended this approach to search for other common functions across all collection files. We then applied specific refactoring methods to each shared function. If the functions were identical across files, we removed the duplicates and created a single function in the library. If modifications existed, we applied case-specific refactoring.

\subsection{Participant-as-Observer Observation Study}

To broaden our study and observations, we adopted a participant-as-observer approach, involving students as free software project newcomers to make practical contributions to the Linux kernel. This approach formed the core activity of a university course, in which we guided students to remove code duplications in the Linux kernel as one possible pathway for contributing to the project in the first phase of the course~[omitted].

The 2025 course offering had 37 students (25 undergraduate and 12 graduate), who were asked to form groups of two or three members, but graduate students could work individually. To assist students in this task, we (a professor and three more experienced graduate students) prepared several alternatives for contributing to the Linux kernel, including simpler contributions such as fixing coding style issues. In this context, we specified two options for removing duplications in the Linux kernel.

\paragraph*{\textbf{Deduplication Option 1}} We ran the tool on the IIO subsystem to generate a list of duplicated function pairs, then curated it to highlight actionable cases for students. Specifically, we (1) kept only pairs within the same file, since cross-driver duplications could involve multiple maintainers or distinct interfaces; (2) removed very short duplications; and (3) manually ranked the remaining pairs, adding annotations to guide students toward entries likely to be accepted. In this option, students were tasked with selecting a recommended entry, devising a way to remove the duplication, and submitting their patches to the IIO mailing list for review.

\paragraph*{\textbf{Deduplication Option 2}} We offered an experience similar to our initial complete participant observation, where students managed the entire workflow: running the tool, analyzing results to identify potential deduplications, creating a patch, and submitting it to the driver's maintainers. This option provided freedom of choice but made the students' task more complex.

Of the students who chose to work on the tool-related tasks, 23 (16 undergraduate and 7 graduate) formed 11 groups to pursue the first alternative. One graduate student opted for the second alternative. Notably, none of these students had prior experience contributing to the Linux kernel, making this their first attempt to submit a patch as newcomers to the project.

It is important to stress that, \textbf{no matter the chosen approach}, we supported all groups by providing review cycles on their contributions before submission, to prepare students for the interactions with the maintainers and alleviate rough, but natural, newcomers' mistakes. After the submission, we helped students understand the feedback from maintainers and gave directions on how to move forward. This close, yet indirect, participation through intimate mentoring characterizes the second phase as our participant-as-observer observation.

For this second phase, our primary data source was the mailing list threads, where all development occurred (available in our Replication Package). Using blog posts, the groups documented their experiences submitting a patch to the Linux kernel~[omitted], and we also leveraged those to complement our observations. % (available in our Replication Package).
%TODO: Add blogs ao pacote de replicação

\section{Results}

In the \textbf{complete participant observation} study (first phase), we chose one function pair found on the AMD Display Driver to approach and apply the systematic approach.
%
We reached a point where we had an idea of removing the duplications found, but we could not complete the refactoring to a state capable of submitting as a patch to the driver. We found that the configuration files on the AMD Display driver are designed to be imported at compilation time using \textit{\#define} macros. The configuration files' design choice makes refactoring duplications on generic approaches trickier, as refactoring code that depends on these files requires significant modification of the configuration files' design and deep knowledge of the codebase, which we, as first-time contributors, did not have. Thus, we opted not to continue investigating this deduplication. We found that the duplicated function from the code reduction existed in five code files, and two other functions were duplicated in those five. All functions across the files were exactly equal, so there was no need to apply the refactoring methods presented in the literature. The refactoring was resumed to create a generic library and fix the compilation targets.

At this stage, we considered the deduplication proposal good enough to submit it to the AMD Display driver as a patch. We sent the first patch on August 9, 2025, and a resend followed on October 9, 2025, due to a lack of initial feedback. Nevertheless, for the first two versions, the maintainers' response requested minor changes for coding style, license use, and alignment with existing conventions. In the third version, the feedback was to move the new generic library functions to an existing file, instead of creating a new one. Then, the fourth version was accepted and integrated into the kernel on February 25, 2025.
%\footnote{V1: \href{https://lore.kernel.org/amd-gfx/20240809210122.23814-1-luanicaro@usp.br/}{https://lore.kernel.org/amd-gfx/20240809210122.23814-1-luanicaro@usp.br/}, V1 (resend): \href{https://lore.kernel.org/amd-gfx/20241009153424.15256-1-luanicaro@usp.br/}{https://lore.kernel.org/amd-gfx/20241009153424.15256-1-luanicaro@usp.br/}, V4 (approved): \href{https://lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br/}{lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br/}}.

Since we suffered a delay in the review process, we directly contacted a maintainer to understand why. From this conversation, we understand that one reason for the delay was that duplicated code enhances the independence of GPU driver code. This approach allows developers to make changes to a specific GPU without needing to test compatibility with others, which helps save a significant amount of time and effort. This example of hardware variations, also mentioned by Kapser and Godfrey~\cite{cloneharm}, leads us to the primary conclusion of our complete participant observation:
%
\begin{highlight-box}{Takeaway T0}
Driver developers do not necessarily view duplicated code negatively, which contradicts standard software engineering practices.
\end{highlight-box}

Building on this insight, we realized that the perception identified in takeaway T0 might not be limited to AMD Display alone. To deepen our understanding, we examined other contributions regarding code duplication reduction and analyzed how different maintainers responded to such patches. This broader analysis aimed to capture diverse maintainer perspectives and contextual factors influencing their decisions.

Table~\ref{tab:stu} summarizes all the contributions sent by the main author and student groups (GID)~\footnote{Since the first author was also a newcomer, the first phase experience was included as Group 0.}. It contains the information about the chosen subsystem (SS) and the similarity threshold (SML). Each entry also has the patch status, takeaways that may have surfaced from the interaction, and a code differential corresponding to the final patch versions before it was dropped or accepted into the code base. The code differential for a set of patches corresponds to the sum of added and removed lines of each patch.

\begin{table}[ht]
\centering
\caption{Summary of newcomer deduplication contributions.}
\begin{tabular}{ |c |c |c |c |c | c| }
\hline
\textbf{GID} & \textbf{SS} & \textbf{SML} & \textbf{Patch Status} & \textbf{Takeaway} & \textbf{Diff} \\
\hline
0 & AMD & 100\% & Accepted (v3) & T0 & +114/-520 \\ \hline
1 & IIO & 100\% & Dropped (v2) & T1, T2 & +7/-14 \\ \hline
2 & IIO & 100\% & Dropped (v1) & T1 & +23/-27 \\ \hline
3 & IIO & 100\% & Accepted (v1) & None & +12/-30 \\ \hline
4 & IIO & 100\% & Accepted (v3) & T1 & +2/-7 \\ \hline
5 & IIO & 90\% & Accepted (v1) & None & +17/-70 \\ \hline
6 & IIO & 90\% & Dropped (v1) & T3 & +95/-92 \\ \hline
7 & IIO & 90\% & Accepted (v3) & None & +20/-36 \\ \hline
8 & IIO & 90\% & Accepted (v4) & T4 & +20/-16 \\ \hline
9 & IIO & 90\% & Accepted (v6) & T4 & +149/-243 \\ \hline
10 & IIO & 90\% & Dropped (v1) & T1, T4 & +32/-58 \\ \hline
11 & IIO & 100\% & Dropped (v1) & None & +2/-12 \\
   & AMD & 90\% & Accepted (v2) & None & +90/-489 \\ \hline

\end{tabular}%

\label{tab:stu}
\end{table}

We derived four takeaways from our \textbf{participant-as-observer observation}. We substantiate these takeaways with quotes from the maintainers that originated during the review process~\footnote{We \textbf{do not claim} these quotes fully represent the maintainers’ stance on code duplication.}.

Groups 1 and 4 worked on similar duplications using the extract method on predicate functions (functions that return a boolean) that, essentially, were the negation of one another. Group 4 initially used macros, but, after feedback from maintainers, both solutions converged toward keeping one of the functions while modifying the other to return the negation of the former. Both cases suffered substantial mutations, even though they were correct in their first iterations, and these changes were primarily driven by the feedback of maintainers regarding readability. For Group 1, a maintainer said that \textit{``[...] the naming as \texttt{\_reg\_check()} is not helpful as it doesn't indicate anything specific is being checked.''}, while for Group 4, the feedback explicitly highlighted the precedence of readability over deduplications: \textit{``I think the old code is more readable than hiding the values in a macro even if it is duplicating a few lines of code.''}.

Even though the deduplication of Group 2 was not similar to that of Group 1 and Group 4 in terms of implementation (it used the parametized method), it also brought forth the readability factor: \textit{``In my view this isn't a significant enough reduction to justify the more complex code.''}.

These experiences revealed a recurring concern among maintainers beyond functional correctness or redundancy reduction. Instead, it reflected a preference for code clarity and comprehension. From these consistent insights that originated from multiple review interactions, we formulated our first takeaway:

\begin{highlight-box}{Takeaway T1 (Readability)}
Maintainers prioritized readability over removing duplications in many different contexts. In their view, going against the \textbf{DRY principle} was worth it if it resulted in easier-to-read and understand code.
\end{highlight-box}

Still focusing on Group 1, maintainers questioned the merit of the proposed patch. While the previous interactions highlighted how maintainers often favor readability over strict code deduplication, this case added a new dimension to our understanding: the \textit{cost-benefit} evaluation of proposed changes. A maintainer pointed out that the effort required to propagate such modifications after acceptance (i.e., during the upstreaming process) could outweigh the perceived benefits: ``\textit{[...] such patches might not worth it since the proposed improvement is very small (and questionable) while the upstreaming process still requires some effort.}''. From this feedback, we crafted our second takeaway:

\begin{highlight-box}{Takeaway T2 (Integration Overhead)}
Maintainers may evaluate the trade-off between the impact of contributions
(not only deduplications) and the struggle to integrate them into the complex
Linux project.
\end{highlight-box}

Groups 3, 6, and 10 used the parameterized method by introducing a new struct used as a parameter to a generic function. Notwithstanding that the same maintainer reviewed all three contributions, only Group 3 had its patch accepted. Our understanding was that this change, in particular, acted on a simple iterative algorithm that (colaterally) also fixed a bug, which favored its straightforward acceptance.

Although removing duplications does not necessarily mean reducing lines of code, the patch from Group 6 added more net lines, which the maintainer pointed out as a drawback. Still, the real reason for the rejection was that the deduplication scattered parts of a complex code that resided close together before: \textit{``Wrapping this up doesn't provide any real advantage, requiring as it does the reviewer to look at this function AND where the value is set rather than seeing them in one place.''}. Needing to jump to a function definition (to see its behavior) or a function call (to see its concrete parameters) in order to understand code is a cognitively demanding task for developers~\cite{questions-programmers-ask}. Merging these informations, we derived our third takeaway:

\begin{highlight-box}{Takeaway T3 (Cognitive Load)}
Maintainers can justify duplications if they avoid having to jump back and forth in the code when parsing through the file.
\end{highlight-box}

The reason for rejecting the patch from Group 10 involved the deduplication lowering readability, as the maintainer was \textit{``[...] not sure the code reduction is sufficiently to cover the resulting loss of readability.''}, reinforcing takeaway T1.

In the interaction of Group 10, there was also a hint that performance could be considered when evaluating the merit of a deduplication. Nonetheless, this notion surfaced aggressively with Group 9, where the contribution could completely remove both duplicated functions using the inline method by leveraging helpers; however, one of the functions had a call that, if eliminated, would result in a significant performance impact: \textit{``Note that is not an appropriate change for the large reads though [...] This is the one case were \texttt{spi\_write\_then\_read()} is probably not appropriate due to the large buffers that are potentially involved.''}. These interactions, coupled with the low-level development context of Linux and driver development, which is usually related to devices with limited resources, produced our fourth and final takeaway.

\begin{highlight-box}{Takeaway T4 (Performance)}
Interactions with maintainers hinted that duplications can be forgiven if they improve performance in contexts where resources are limited.
\end{highlight-box}

Groups 7 and 8 used the extract method using helpers to remove duplicated code in the function's body. Group 7 refined its patch from the maintainers' feedback, resulting in a smooth acceptance process that did not produce any insights. Group 8 followed a similar course, but generated an interaction reinforcing takeaway T4. In the feedback of version 1, a maintainer claimed that \textit{``Even though there is less code repetition, we now have an extra comparison [...]''}, with another maintainer corroborating this in version three by saying \textit{``There is always a balance/trade-off between modularity and execution speed. I agree with \texttt{anonymous-maintainer}'s reply in the first patch [...]''}, regardless of the fact that the patch was eventually accepted in version four.

We did not explore the contributions of Group 5 and Group 11 in detail, as they did not result in any insights regarding these conflicting perspectives on software engineering practices. However, they were still meaningful as they produced impactful contributions that reduced a lot of duplicated code. Especially in Group 11, where a single student executed the whole workflow we did in the first phase, and managed to remove four hundred net lines of code.

In this sense, we discuss another aspect of importance stemming from this work: using the developed tool to create deduplication opportunities in the Linux kernel can be an excellent introduction for newcomers.

To illustrate this point, we display some statistics concerning the 13 contributions contemplated in this paper:

\begin{itemize}
  \item Collectively, the patches proposed the insertion of 585 lines and the removal of 1626 lines.
  \item 8 out of 13 contributions were accepted, resulting in a success rate of 62\%.
  \item For accepted contributions, the code diff value was +255/-1152, averaging +31.9/-144.0 additions/removals.
  \item For rejected contributions, the code diff value was +181/-231, averaging +36.2/-46.2 additions/removals.
\end{itemize}

\section{Concluding Remarks}

This ethnographic study, combining complete participant and participant-as-observer observations, provided a realistic view of the opportunities and challenges in removing code duplications detected by the tool. First-time Linux kernel contributors could effectively use the tool to identify duplicated functions and submit meaningful patches.

Throughout this process, we observed that maintainers do not necessarily regard code duplication as inherently harmful. Instead, their decisions reflect a nuanced and context-dependent understanding of software quality. As evidenced by our takeaways\footnote{It is paramount to state that we \textbf{are not affirming} that these quotes represent the complete positioning of the maintainers who wrote them on code duplication.}, readability often outweighed strict adherence to the DRY principle (T1), while the effort required to propagate changes also influenced maintainers’ acceptance decisions (T2). Moreover, avoiding cognitive overhead when navigating the code (T3) and preserving performance in resource-constrained environments (T4) emerged as additional justifications for tolerating certain forms of duplication.

Overall, these findings reveal that, within the Linux kernel community, the evaluation of code duplication extends beyond technical redundancy. Maintainability, clarity, and practical trade-offs play a decisive role in shaping maintainers’ judgments, suggesting that DRY is sometimes less valuable than keeping the code understandable, efficient, and integrated smoothly into a large and evolving system.

\section*{Replication Package}

All anonymized threads containing the entire review process and interactions with maintainers are available in this Zenodo repository: \href{https://doi.org/10.5281/zenodo.17432193}{https://doi.org/10.5281/zenodo.17432193}.

To avoid violating the double-anonymized review process, the anonymous repository of the developed tool is available at \url{https://anonymous.4open.science/r/arkanjo-C548}. The tool website link and omitted references will be added in the camera-ready version of the paper.

\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{IEEEabrv,references}

\end{document}